R Notebook

### 2016-09-30 manuscript/draft.Rmd

Quality of predictions from SDMs is important:
  * helps determine policy/set priorities
  * something something climate change
  * need to know both accuracy of point estimates and how reasonable our prediction intervals are.
  * Overconfident predictions can be worse than uncertainty
  * Quality of prediction is one metric for how much we know about ecological processes
SDM assumptions:
  * we've measured all the important variables (sites don't have any other important properties)
  * Something like stationarity:
    * e.g. "systems are at equilibrium now and will be at equilibrium in future"
    * or "species track predictor variables quickly" (no lags)
    * etc
  * independent/identically distributed observations across species, years
  * Space substitutes well for time
Difficult to test SDM assumptions/predictions because of time scales
  * Most scientists making predictions today will have retired by 2050
  * Use long-term data sets such as BBS to evaluate predictive skill
    * Note the large change in average temperatures since the 80's; comparable to the changes expected in the next N years
  * Test with simple baselines (e.g. random walk, future-is-like-present)
  * Focus on richness because... reasons.

@@

* BBS + retriever
* Monthly climate data from interpolated weather stations
* NDVI
* Altitude

@@

* Temporal richness variability doesn't track the predictors we use; sites have "inertia"
* More important to use a reasonable error model than to get the best predictors \& basis expansions
* We've posted predictions for the next N years of BBS data, using several methods, and invite others to do so as well.
* In many ways, this is the best-case scenario for SDMs:
  * large data set, known \& consistent sampling technique over a period of decades
  * True absences
  * extended predictor set
  * birds arguably migrate quickly enough to make equlibrium \& especially space-for-time assumptions less ridiculous
  * errors for individual species can cancel out; only asking for aggregate values
* Value of stacking versus predicting the quantitity of interest directly
*

>>>>

Quality of predictions is important:
  * helps determine policy/set priorities
  * something something climate change
  * need to know both accuracy of point estimates and how reasonable our prediction intervals are.
  * Overconfident predictions can be worse than uncertainty
  * Quality of prediction is one metric for how much we know about ecological processes
Dominant method is stacked SDMs:
  *
Difficult to test model assumptions/predictions because of time scales
  * Most scientists making predictions today will have retired by 2050
  * Use long-term data sets such as BBS to evaluate predictive skill
    * Note the large change in average temperatures since the 80's; comparable to the changes expected in the next N years
  * Test with simple baselines (e.g. random walk, future-is-like-present)
  * Focus on richness/diversity because \[citations saying that it's important\]

@@

* BBS + retriever
* Monthly climate data from interpolated weather stations (PRISM)
* NDVI (GIMMS)
* Altitude (SRTM 90m Digital Elevation Database via raster::getData)

@@

* Temporal richness variability doesn't track the predictors we use; sites have "inertia"
* More important to use a reasonable error model than to get the best predictors \& basis expansions
* We've posted predictions for the next N years of BBS data, using several methods, and invite others to do so as well.
* In many ways, this is the best-case scenario for SDMs:
  * large data set, known \& consistent sampling technique over a period of decades
  * True absences
  * extended predictor set
  * birds arguably migrate quickly enough to make equlibrium \& especially space-for-time assumptions less ridiculous
  * errors for individual species can cancel out; only asking for aggregate values
* Value of stacking versus predicting the quantitity of interest directly

## misc

SDM assumptions:
  * we've measured all the important variables (sites don't have any other important properties)
  * Something like stationarity:
    * e.g. "systems are at equilibrium now and will be at equilibrium in future"
    * or "species track predictor variables quickly" (no lags)
    * etc
  * independent/identically distributed observations across species, years
  * Space substitutes well for time
